---
title: "ds202_hw5"
author: "YangfanCai/yangfanc"
date: "3/31/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1.The complete dataset is too large. Download and import the alcohol sales data for Ames only:
a.Go to https://data.iowa.gov/Sales-Distribution/2019-Iowa-Liquor-Sales/38x4-vs5h
b.Click Filter – Filter – Go advanced now
c.Add a New Filter Condition: City is Ames.
d.Export – CSV – Download
e.Use readr::read_csv to import.
```{r}
library(dplyr)
library(ggplot2)
library(lubridate)
library(tidyverse)
if (!require("tidyverse")) {
install.packages("tidyverse")
}
if (!require("dplyr")) {
install.packages("dplyr")
}
if (!require("ggplot2")) {
install.packages("ggplot2")
}
if (!require("readr")) {
install.packages("readr")
}
if (!require("lubridate")) {
install.packages("lubridate")
}
```


```{r}
ames <- readr::read_csv('~/Downloads/2019_Iowa_Liquor_Sales.csv')

```


# 2.Data cleaning:
a. extract geographic latitude and longitude from the variable Store Location
b. make sure the variables are in the correct type, by converting incorrect variable types.
c. extract year, month and day from the variable Date as separate columns.
```{r}
#a
amescleana <- ames %>%
  separate('Store Location', c("Address", "City", "Coordinates"), sep = "\n") %>%
  mutate(Coordinates = gsub("[()]", "", Coordinates)) %>%
  separate("Coordinates", c("Latitude", "Longitude"), sep = ",")
amescleana
```


```{r}
amesclean1 <- ames %>% 
  separate(`Store Location`, c('Point', 'long', 'lat'), sep=' ') %>% 
  separate(long, c('extra', 'long'), sep = '\\(') %>% 
  separate(lat, c('lat', 'extra2'), sep = '\\)') %>% 
  select(-extra, -extra2, -Point)
amesclean1
```






```{r}
#b
amesb <- ames %>% 
  separate(`Store Location`, c('Point', 'long', 'lat'), sep=' ') %>% 
  separate(long, c('extra', 'long'), sep = '\\(') %>% 
  separate(lat, c('lat', 'extra2'), sep = '\\)') %>% 
  select(-extra, -extra2, -Point)
amesb
```


```{r}
ames2 <- amesclean1 %>% 
  mutate(long = as.numeric(long),
         lat = as.numeric(lat))
ames2
```



```{r}
#c
ames3 <- ames2 %>% 
  separate(Date, into = c('Month','Day','Year'), sep = '/') %>% 
  mutate(Month = as.integer(Month),
         Day = as.integer(Day),
         Year = as.integer(Year))
ames3
```


# 3.First overview:
a.Plot a scatterplot of lat and long of store locations.

b.Provide a visual breakdown of the liquor category (by Category Name). Include volume sold in the breakdown.


```{r}
#a
ames3 %>% ggplot(aes(x = long, y = lat))+
  geom_point(na.rm = TRUE) + geom_jitter(na.rm = TRUE)+
  ggtitle("Scatterplot of Latitude Vs Longitude")+ xlab("Longitude")+ ylab("Latitude") 
```






```{r}
#b
ames3 %>% ggplot(aes(x = `Category Name`, y = `Volume Sold (Liters)`))+
  geom_line(aes(color = `Category Name`))+ theme(legend.position = "none", axis.text.x = element_text(angle = 45))+ coord_flip()+ 
  ggtitle("Amount_of_Alcohol_Sold_by_Category_Name")+ xlab("Category_Name")+ ylab("Amount_in_Liters_Sold") 
```


# 4. Principal Component Analysis for Functional Data on Riemannian Manifolds and Spheres

```{r}
ames4 <- ames3 %>% 
  group_by(Year, Month, Day) %>% 
  summarise(`Total Sales`=n(),
            `Volume`  = sum(`Volume Sold (Gallons)`),
            `Money Spent` = sum(`Sale (Dollars)`))
ames4
```

```{r}
ames3 %>%
  group_by(`Bottles Sold`) %>%
  ggplot(aes(x=Day, y=`Volume Sold (Liters)`)) + geom_point() +facet_wrap(~Month)
q4 <- ames3 %>% filter(Year == 2019)
q41 <- q4%>% group_by(Day,Month) %>% 
  summarise(sumBottleSold = sum(`Bottles Sold`), sumSale = sum(`Sale (Dollars)`),
                                sumVolumeSold = sum(`Volume Sold (Liters)`)) 
q41
```



```{r}
amesF  <- ames3 %>% group_by(Month,Day) %>% summarize(Number=n(),Quantity=sum(`Volume Sold (Gallons)`),Revenue = sum(`Sale (Dollars)`))
summary(amesF)
```






# 5.Plot volume sold by day (use a scatterplot of volume by day and facet by month). Discover any trend with the help of geom_smooth. Describe any patterns.
```{r}
ames4 %>% 
  ggplot(aes(x=Day, y=Volume))+geom_point()+geom_smooth(se = FALSE)+facet_wrap(~Month)+theme_bw()
```


```{r}
amesF %>% ggplot(aes(x=Day,y=Quantity))+geom_jitter()+geom_smooth(method="lm")+facet_wrap(~Month)
```





# 6.Find the dates for ISU football home games in Fall 2019. Can you see a pattern?
```{r}
ames4 %>% 
  filter(Month %in% c(8,9,10,11)) %>% 
  ggplot(aes(x=Day, y=Volume))+geom_point()+geom_smooth(se = FALSE)+facet_wrap(~Month)+theme_bw()
```

```{r}
ggplot(ames3, aes(x=Day,y=`Volume Sold (Liters)`)) +
  geom_col() +
  facet_wrap(vars(Month)) + 
  theme(axis.text.x = element_blank()) +
  labs(title = "Volume_Sold_by_Day_faceted_by_Month")
```


The dates for the football season were 8/31, 9/14, 9/21, 9/28, 10/5, 10/12, 10/19, 10/26, 11/9, 11/16, 11/23, and 11/30.There are always big increases in sales the few days prior to home football games.
